---
title: "DATA607_Project 3 - Data Science Skills"
author: "Team 3: David Simbandumwe, Thomas Buonora, Charles Ugiagbe, Jaya Veluri"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  rmdformats::robobook:
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(knitr)
library(XML)
library(curl)
library(RCurl)
library(jsonlite)
library(dplyr)
library(openintro)
library(RMySQL)
library(ggplot2)
library(stringr)
library(stringi)

library(Hmisc)
library(corrplot)

```



# Introduction





```{r}

usr <- keyring::key_list("DATA607")[1,2]
pwd <-  keyring::key_get("DATA607", usr)
con = dbConnect(MySQL(), user=usr, password=pwd, dbname='DATA607', host='localhost')


rs = dbSendQuery(con, "select * 
            from SkillsMeta")
ds_skills_list_df = fetch(rs, n=-1)


dbDisconnect(con)


```





# Read data
```{r}

# read skills data from csv
skills_df <- read_csv( file = "/Users/dsimbandumwe/dev/cuny/data_607_T3/DATA607Team3/source/Final_Train_Dataset.csv")
names(skills_df)[1] <- "id"


# build a temporary data frame 
tmp <- skills_df %>%
    select(id, key_skills) %>%
    separate_rows(
        key_skills,
        convert = TRUE,
        sep = "\\,"
    )
    

tmp <- tmp %>% 
    mutate(
        key_skills = str_to_lower(key_skills),
        key_skills = str_replace_all(key_skills, "\\.{3}", "" ),
        key_skills = str_trim(key_skills)
    )

tmp <- tmp %>% 
  right_join(ds_skills_list_df, by="key_skills" ) %>%
  rename(
      key_skills_id = id.y,
      id = id.x
  )

# build a list of user ids that have data science skills
id_df <- tmp %>%
    select(id) %>%
    distinct()


# join the temporary dataframe with the original dataframe
skills_df <- skills_df %>% 
    inner_join( id_df,by="id" ) %>%
    right_join(tmp, by="id") %>%
    select(-c(job_description,job_desig,key_skills.x)) %>%
    rename(
      key_skills = key_skills.y
    )


# update the salary
skills_df <- skills_df %>%
    separate(
        salary,
        c("min_salary" , "max_salary"),
        convert = TRUE,
        sep = "to"
    )


#write out csv file
write.csv(skills_df, "/Users/dsimbandumwe/dev/cuny/data_607_T3/DATA607Team3/output/skillsOutput.csv", row.names=FALSE)


```




```{r}

emp_df <-  read.csv(
  "/Users/dsimbandumwe/dev/cuny/data_607_T3/DATA607Team3/source/multipleChoiceResponses.csv", 
  header=T,sep="," 
)

tmp <- emp_df %>% select(Q1, Q2, Q3, Q4, Q6, Q8, Q9, starts_with("Q13"), starts_with("Q16"))
tmp <- tmp %>% select(-c("Q13_OTHER_TEXT","Q16_OTHER_TEXT"))

# t3 <- tmp %>% group_by(Q6) %>% 
#   mutate(
#    n = n() 
#   ) %>%
#   select(Q6,n) %>% 
#   distinct()

tmp <- tmp %>%
  filter(grepl("Data Scientist",Q6)) %>%
  mutate (
     id = row_number()
  )
 

tmp <- slice(tmp,-(1:1))
tmp <- tmp %>% pivot_longer(
                starts_with("Q13") | starts_with("Q16"),
                names_to = "q",
                values_to = "ans"
              ) 

tmp <- tmp %>% 
  mutate (
    ans = str_squish(ans),
    ans = str_to_lower(ans)
  )

tmp <- tmp %>% 
  right_join(ds_skills_list_df, by=c("ans" = "key_skills") ) %>%
  rename(
      key_skills_id = id.y,
      id = id.x
  )


tmp <- tmp %>% 
  filter (ans != "" & Q9 != "") %>%
  filter(!grepl("I do not",Q9)) %>%
  mutate (
    Q9 = str_replace(Q9, "\\+",""),
    Q9 = str_replace(Q9, ",000","")
  )

  
tmp <- tmp %>% 
  separate(
      Q9,
      c("min_salary" , "max_salary"),
      convert = TRUE,
      sep = "-"
  ) 


tmp <- tmp %>%
  transform(
    min_salary = as.numeric(min_salary),
    max_salary = as.numeric(max_salary)
  ) 


emp_df <- tmp %>%
    rename(
      gender = Q1,
      age = Q2,
      location = Q3,
      education = Q4,
      title = Q6,
      experience = Q8
    ) %>%
  select (id, q, ans, key_skills_id, min_salary, max_salary, gender, age, location, education, title, experience)


write.csv(emp_df,"/Users/dsimbandumwe/dev/cuny/data_607_T3/DATA607Team3/output/multipleChoiceOutput.csv")

```










# Analysis Open Roles


```{r}

s1 <- skills_df %>% group_by(key_skills) %>%
  mutate(
    n = n()
  ) %>%
  select(key_skills, n) %>%
  distinct()

s1 %>% 
  ggplot(aes( y=key_skills, x=n)) +
  geom_bar(position="dodge", stat="identity") + 
  labs(title = "Instances of Specific Skills in the Dataset" )

```




```{r}

skills_df %>% 
  ggplot() +
  geom_boxplot(mapping = aes(y=key_skills, x=max_salary, color=key_skills)) + 
  labs (title = "Mapping Salary to Skillset" )


```




```{r}

# create a wide dataframe for correlation
t <- skills_df %>% 
  mutate(
    flag = 100
  ) %>%
  select(-c(key_skills_id,job_type)) %>%
  mutate (
   key_skills = str_replace_all(key_skills, " ", "_"),
   key_skills = str_squish(key_skills),
   row = row_number()
  ) 
t <- t %>%
  pivot_wider(
    names_from = key_skills,
    values_from = flag,
    values_fill = 0
  ) 
t <- t %>% select(-c(id, experience, location, row, company_name_encoded)) 


t.rcorr = rcorr(as.matrix(t))
t.rcorr
    
    
t_cor = cor(t, method = c("spearman"))
corrplot(t_cor, title="Coorelation Income By Skills")

```




# Analytis Job Seekers

```{r}

s1 <- emp_df %>% group_by(ans) %>%
  mutate(
    n = n()
  ) %>%
  select(ans, n) %>%
  distinct()

s1 %>% 
  ggplot(aes( y=ans, x=n)) +
  geom_bar(position="dodge", stat="identity") + 
  labs(title = "Instances of Specific Skills in the Dataset" )

```


```{r}
emp_df %>% 
  ggplot() +
  geom_boxplot(mapping = aes(y=ans, x=max_salary, color=ans)) + 
  labs (title = "Mapping Salary to Skillset" )
```




```{r}

# create a wide dataframe for correlation
t2 <- emp_df %>% 
  mutate(
    flag = 100
  ) %>%
  mutate (
   ans = str_replace_all(ans, " ", "_"),
   ans = str_squish(ans),
   row = row_number()
  ) 
t2 <- t2 %>%
  pivot_wider(
    names_from = ans,
    values_from = flag,
    values_fill = 0
  ) 
t2 <- t2 %>% select(-c(id, q, key_skills_id ,experience, title, age, gender, location, education )) 
t2 <- t2 %>% drop_na()


t2.rcorr = rcorr(as.matrix(t2))
t2.rcorr
    
    
t2_cor = cor(t, method = c("spearman"))
corrplot(t_cor, title="Coorelation Income By Skills (Job Seeker)")

```



# Conclusions

Based on our analysis we can identify a few skills that do not correlate with income or salary. However given how low the overal levels of correlation it is difficult to come to any additional conclusions.  

- **Open Roles**
- marketing automation 0.2043 min and 0.1458 max
- data privacy 0.4249 min and 0.3341 max
- matlab 0.7624 min and 0.6057 max
- big data analytics min 0.7172 and 0.1767 max
- and interestingly enough pyhton has a slightly negative correlation


- **Job Seekers**
- python 0.3131 min and 0.3132 max

The results could be a factor of our limited datasets or it could also be caused by additional factors that impact salary that are not included in the data. Some items could include:
- where you received your college education
- industry
- geography within the course grained location (NY, SFO markets)










